---
title: "A portfolio website by Philip Fosser"
execute: 
  echo: false
---
------------------------------------------------------------------------

```{r}
#| warning: false
library(tidyverse)
library(htmltools)
```

### An introduction.

Hi there! I'm Philip, and welcome to what serves as my portfolio website, though you probably guessed what this website is based on the title already. If you're here, you're likely a recruiter or someone entrusted with the important task of vetting potential candidates. If that is the case, I hope you find this website to serve is intended purpose. To showcase that I can use code to analyse data, document, visualize results etc. And that the website proves that I'm not full of it when I make claims about my abilities, or that I use some of my spare time to improve upon my data analysis and coding abilities, as i claim in my resume. 

However, if you happen to be my wonderful girlfriend browsing through, I created the sleep analysis section specifically for you, so please consider it my loving reminder of just how important sleep really is.

------------------------------------------------------------------------

### About me (personal life)
::: column-page
```{r}

x <- tibble(
  Heading = c(
    "Daily learning", 
    "Exercise",
    "Small business",
    "Cooking and baking", 
    "Traveling"
  ),
  Description = c(
    'Every day, I carve out time to stay curious and informed???diving briefly into major global news that moves markets, and spending the bulk of my attention on practical insights that I can immediately apply. After all, it has been proven that people learn best when new information is immideatly taken into use, case in point, this segmented portion of the introduciton.',
    
    "I've been consistently doing stregnth training since I was 16, and it has fundementally shaped me in more ways than one. Beyond building resilience and confidence, both game-changers for overall well-being, it even fixed the chronic back pain from my naturally crooked spine. Now the only effect of my somewhat crooked spine, is that it makes me a bit more vertically challenged than i need to be.",
    
    "In 2020 I started a small, hobby woodworking business that I still operate today (It's not a registered business, and I make sure to make less than what makes me liable to pay taxes on it). I started this business after I had decided that I wanted to fix one of my gratest weaknesses. My complete incopetence when it came to doing anything practical with my hands.",
    
    "Try as one might, consistently training and learning is a hopeless endeavour if you don't feed yourself properly. So, I have made an earnest attempt at cooking delicious and healthy food to make sure I feel as good as possible. I prefer baking to cooking, but the results of my baking does little to improve my health, and is in fact more of a detriment to it. Tastes good dough.",
    
    "To say one likes to travel, or that it is a hobby, is a very clever way to say one enjoys going on vacation. It's like saying that you like eating good food. But I do like eating good food, and I do enjoy to travel. One of my dreams is to take a cross country, or cross continent road trip on a motorcycle. Now I only need to learn how to ride a motorcycle."
    
  )
) %>% 
  mutate(
    Heading =  map(
      Heading, 
      \(x) htmltools::div(htmltools::strong(x))
    ),
      Description =  map(
      Description, 
      htmltools::div
    ),
    combined = map2(
      Heading,
      Description,
      \(x,y) htmltools::div(
        style = htmltools::css(
          width = "300px",
          margin_right = "20px",
          margin_bottom = "20px"
          ),
          x,
          y
        )
      )
    )
  
htmltools::div(
  style = css(
    display = "flex",
    flex_wrap = "wrap",
  ),
  x$combined
)
```
:::

---

### What can I do?

Perhaps the most important question I can ask myself is; why should anyone choose to hire me?

I have to admit, figuring out a clever answer to that question that would encompasses all the possible needs of any of the positions that I have applied to, is a bit outside my abilities. So you'll have to excuse my verbose answer, as I make an attempt at advocating for myself. 

The person reading, is probably already familiar with my resume, as it's where you got the link to this website. So I will not bore you with my formal education, work experience etc. The argument I use for why I would be a good asset to any workplace, isn't tied to ***[what]{style="text-decoration: underline;"}*** my abilities or achievements are, but rather ***[how]{style="text-decoration: underline;"}***. 
As for the ***[how]{style="text-decoration: underline;"}*** I got them by being exceptionally dedicated to achieving my goals. That is how I, in four years, managed to both start and finish a bachelors and a masters degree in something that did not come naturally to me (and managing to achieve a B average on top of that). And that is also how I in that same timeperiod learned woodworking, coding in two different languages, restored my mothers house, saved up to purchase realestate and created an incredibly healthy relationship with my girlfriend, all while getting in better and better shape by each year that came and went.

I will bring that same dedication to any workplace, as being among the top of my field is a part of that goal. It's this trait that I believe will bring high monetary and strategic value to any workplace that I'm a part of. 

---

### Contact information

I can be reached by email at [phankrfo@gmail.com](mailto:phankrfo@gmail.com),
or by phone: [97989947](tel:97989947)